The Minimal Retroreflective Microfacet Model

### 2025-12-19 article.tex
Fix placements.

\usepackage{amsmath}
\usepackage{listings}
\usepackage{xcolor}
\usepackage[normalem]{ulem}
\usepackage{siunitx}
\usepackage{csquotes}
\usepackage{url}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{mathtools}
\usepackage{xfrac}
\usepackage{comment}
\usepackage{natbib}

@@

\teaser{
  \vspace*{-0.5cm}
    \hfill
    \includegraphics[width=0.25\linewidth]{figures/cone/nonretro_cone_cropped_label.png}
    \includegraphics[width=0.25\linewidth]{figures/jacket/nonretro_jacket_cropped_label.png}
    \includegraphics[width=0.25\linewidth]{figures/shoe/nonretro_shoe_cropped_label.png} \\
    \vspace{0.06cm}
    \includegraphics[width=0.25\linewidth]{figures/cone/retro_cone_cropped_label.png}
    \includegraphics[width=0.25\linewidth]{figures/jacket/retro_jacket_cropped_label.png}
    \includegraphics[width=0.25\linewidth]{figures/shoe/retro_shoe_cropped_label.png}
    \hfill
  \caption{Comparison of our ``Minimal Retroreflective Microfacet'' (MRM) retroreflector model (bottom row) obtained by an almost trivial modification of the equivalent roughness GGX microfacet model (top row). The lighting in all cases is aligned with the view direction, exhibiting a strong retro-reflective peak.}
   \vspace{-0.5cm}
  \label{fig:teaser}
}

@@

Given an implementation of a regular microfacet BSDF, extending it to retro-reflection is extremely straightforward:

\begin{itemize}
\item
  Evaluation merely needs to replace $\vecv$ with $\vecvv$ upfront.
\item
  Similarly, importance sampling of $\vecl$ given $\vecv$ can be realized by replacing $\vecv$ with $\vecvv$ upfront and then importance sampling the regular microfacet BSDF.
  This may include low variance sampling using the domain of visible microfacets \cite{HeitzIS}.
\item
  As the albedos of standard BSDF and retro-reflective BSDF are identical, compensating for energy loss in the sense of \cite{KelemenBRDF} can be realized using the same data tables.
\item Pseudo-code is provided in Listing~\ref{lst:implementation}.
\end{itemize}

\noindent
\begin{lstlisting}[
    style=snippet,
    language=GLSL,
    caption=Pseudo code to implement the retro-reflective microfacet model using existing regular microfacet BSDF implementation,
    label=lst:implementation,
    nolol=true,
    frame=trBL
    ]
Sample_result sample_bsdf_retro(vec3 V, vec3 L, vec3 N, Bsdf_params params)
{
    vec3 Vp = -V + 2.0 * dot(V, N) * N;
    return sample_bsdf(V, L, N, params);
}
Eval_result eval_bsdf_retro(vec3 V, vec3 L, vec3 N, Bsdf_params params)
{
    vec3 Vp = -V + 2.0 * dot(V, N) * N;
    return eval_bsdf(V, L, N, params);
}
\end{lstlisting}

@@

\begin{figure}[!b]
    \centering
    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/classic_angle_0_cropped.png}
    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/classic_angle_45_cropped.png}
    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/classic_angle_90_cropped.png} \\
    \vspace{0.06cm}
    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/retro_angle_0_cropped.png}
    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/retro_angle_45_cropped.png}
    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/retro_angle_90_cropped.png}
  \caption{Renders of a metallic shaderball with either a classic GGX (top row) or retro-reflective MRM material (bottom row) applied, for various camera angles and a fixed light direction. The relative angle between the view and light direction is $0^\circ$ (left), $45^\circ$ (middle), and $90^\circ$ (right).}
  \label{fig:conductor_renders}
\end{figure}

@@

Figure~\ref{fig:dielectric_renders} demonstrates this in renders of a dielectric shaderball, where the IOR of the dielectric material is either $1.0$ (top row) or $1.5$ (bottom row).
Left to right we show three combinations of BRDF and BTDF: the classic microfacet BSDF $f_r$ and BTDF $f_t$ (left), the retro-reflective BRDF $f_{r,\text{retro}}$ plus retro-reflective BTDF $f_{t,\text{retro}}$ (middle), and retro-reflective BRDF $f_{r,\text{retro}}$ combined with the classic (non-retroreflective) BTDF $f_t$ (right).
\begin{figure}[!tb]
    \centering
    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/classic_ior_1_cropped_label.png }
    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroRT_ior_1_cropped_label.png }
    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroR_ior_1_cropped_label.png } \\
    \vspace{0.06cm}
    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/classic_ior_1.5_cropped_label.png }
    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroRT_ior_1.5_cropped_label.png}
    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroR_ior_1.5_cropped_label.png }
  \caption{Renders of a glass shaderball illuminated from the side. The index of refraction is $1.0$ (top) or $1.5$ (bottom). From left to right: classic microfacet BSDF (left), retro-reflective BSDF (middle), and combined retro-reflective BRDF and classic BTDF (right).  }
  \label{fig:dielectric_renders}
\end{figure}

>>>>

\usepackage{amsmath}
\usepackage{listings}
\usepackage{xcolor}
\usepackage[normalem]{ulem}
\usepackage{siunitx}
\usepackage{csquotes}
\usepackage{url}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{mathtools}
\usepackage{xfrac}
\usepackage{comment}
\usepackage{natbib}
\usepackage[font={small,it}]{caption}

@@

\teaser{
  \vspace*{-0.5cm}
    \includegraphics[width=0.3\linewidth]{figures/cone/nonretro_cone_cropped_label.png}
    \includegraphics[width=0.3\linewidth]{figures/jacket/nonretro_jacket_cropped_label.png}
    \includegraphics[width=0.3\linewidth]{figures/shoe/nonretro_shoe_cropped_label.png} \\
    \vspace{0.06cm}
    \includegraphics[width=0.3\linewidth]{figures/cone/retro_cone_cropped_label.png}
    \includegraphics[width=0.3\linewidth]{figures/jacket/retro_jacket_cropped_label.png}
    \includegraphics[width=0.3\linewidth]{figures/shoe/retro_shoe_cropped_label.png}
  \caption{Comparison of our ``Minimal Retroreflective Microfacet'' (MRM) retroreflector model (bottom row) obtained by an almost trivial modification of the equivalent roughness GGX microfacet model (top row). The lighting in all cases is aligned with the view direction, exhibiting a strong retro-reflective peak.}
   \vspace{-0.5cm}
  \label{fig:teaser}
}

@@

Given an implementation of a regular microfacet BSDF, extending it to retro-reflection is extremely straightforward (see the pseudo-code provided in Listing~\ref{lst:implementation}):

\begin{itemize}
\item
  Evaluation merely needs to replace $\vecv$ with $\vecvv$ upfront.
\item
  Similarly, importance sampling of $\vecl$ given $\vecv$ can be realized by replacing $\vecv$ with $\vecvv$ upfront and then importance sampling the regular microfacet BSDF.
  This may include low variance sampling using the domain of visible microfacets \cite{HeitzIS}.
\item
  As the albedos of standard BSDF and retro-reflective BSDF are identical, compensating for energy loss in the sense of \cite{KelemenBRDF} can be realized using the same data tables.
\end{itemize}

\noindent
\begin{lstlisting}[
    style=snippet,
    language=GLSL,
    caption=Pseudo code to implement the retro-reflective microfacet model using an existing regular microfacet BSDF implementation,
    label=lst:implementation,
    nolol=true,
    frame=trBL
    ]
Sample_result sample_bsdf_retro(vec3 V, vec3 L, vec3 N, Bsdf_params params)
{
    vec3 Vp = -V + 2.0 * dot(V, N) * N;
    return sample_bsdf(V, L, N, params);
}
Eval_result eval_bsdf_retro(vec3 V, vec3 L, vec3 N, Bsdf_params params)
{
    vec3 Vp = -V + 2.0 * dot(V, N) * N;
    return eval_bsdf(V, L, N, params);
}
\end{lstlisting}

@@

@@

Figure~\ref{fig:dielectric_renders} demonstrates this in renders of a dielectric shaderball, where the IOR of the dielectric material is either $1.0$ (top row) or $1.5$ (bottom row).
Left to right we show three combinations of BRDF and BTDF: the classic microfacet BSDF $f_r$ and BTDF $f_t$ (left), the retro-reflective BRDF $f_{r,\text{retro}}$ plus retro-reflective BTDF $f_{t,\text{retro}}$ (middle), and retro-reflective BRDF $f_{r,\text{retro}}$ combined with the classic (non-retroreflective) BTDF $f_t$ (right).

\begin{figure}[!ht]
    \centering
    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/classic_angle_0_cropped.png}
    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/classic_angle_45_cropped.png}
    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/classic_angle_90_cropped.png} \\
    \vspace{0.06cm}
    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/retro_angle_0_cropped.png}
    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/retro_angle_45_cropped.png}
    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/retro_angle_90_cropped.png}
  \captionsetup{font=small}
  \caption{Renders of a metallic shaderball with either a classic GGX (top row) or retro-reflective MRM material (bottom row) applied, for various camera angles and a fixed light direction. The relative angle between the view and light direction is $0^\circ$ (left), $45^\circ$ (middle), and $90^\circ$ (right). \vspace{-0.5cm}}
  \label{fig:conductor_renders}
\end{figure}

\begin{figure}[!hb]
    \centering
    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/classic_ior_1_cropped_label.png }
    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroRT_ior_1_cropped_label.png }
    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroR_ior_1_cropped_label.png } \\
    \vspace{0.06cm}
    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/classic_ior_1.5_cropped_label.png }
    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroRT_ior_1.5_cropped_label.png}
    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroR_ior_1.5_cropped_label.png }
  \caption{Renders of a glass shaderball illuminated from the side. The index of refraction is $1.0$ (top) or $1.5$ (bottom). From left to right: classic microfacet BSDF (left), retro-reflective BSDF (middle), and combined retro-reflective BRDF and classic BTDF (right).  }
  \label{fig:dielectric_renders}
\end{figure}
